---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# workspace

`workspace` is an R package designed to simplify the creation and management of workspaces. Its main goal is to provide a standardized solution for organizing and sharing data in applications that require workspace management.

This package is aimed at developers and analysts seeking to standardize data management in multi-environment or cross-application contexts. With `workspace`, creating, manipulating, and sharing organized archives becomes straightforward.

## Key features

- Generation of structured archives:
  - Supports adding datasets in parquet format.
  - Enables storing JSON files for semi-structured data.
  - Handles R objects stored in RDS format.

- Enhanced interoperability:
  - Provides a well-defined and consistent structure for organizing data.
  - Facilitates data exchange and processing across different applications or environments.

- Offers tools to easily read and store data within workspaces.


## Installation

You can install the development version of workspace from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("ardata-fr/workspace")
```

## Example of workspace creation

This is a basic example which shows you how to create a workspace:

```{r}
library(workspace)
z <- new_workspace()

# store datasets
z <- store_dataset(x = z, dataset = iris, name = "iris_dataset")
z <- store_dataset(x = z, dataset = mtcars, name = "mtcars")

# store json
json_str <- paste0("{\"first_name\": \"John\",\"last_name\": \"Smith\",\"is_alive\": true,",
"\"age\": 27, \"address\": { \"street_address\": \"21 2nd Street\",",
"\"city\": \"New York\",\"state\": \"NY\",\"postal_code\": \"10021-3100\"",
"}}")
z <- store_json(
  x = z,
  name = "json-example",
  json_str = json_str,
  filename = "json-example.json",
  timestamp = "2023-11-12 11:37:41",
  subdir = "blah"
)

# pack workspace as a zip to share it
workspace_zip_file <- tempfile(fileext = ".zip")
pack_workspace(x = z, file = workspace_zip_file)
```

## Example of workspace data extraction

This is a basic example which shows you how to extract data from a workspace:

```{r}
z <- unpack_workspace(file = workspace_zip_file)
list_object_in_workspace(z)

dataset <- read_dataset_in_workspace(z, name = "mtcars")
print(head(dataset))

# store json
json_str <- read_json_str_in_workspace(z, name = "json-example", subdir = "blah")
substr(json_str, 1, 20) |> print()
```

## Example Shiny app using workspace

`workspace` can be useful for managing datasets in Shiny apps. The example shows how to create, add datasets, pack and unpack a workspace. 

```{r, eval = FALSE}
library(workspace)
library(shiny)
library(readr)
library(shinyWidgets)

# Define UI for application that draws a histogram
ui <- fluidPage(
  tags$h4("Add datasets to workspace"),
  shiny::fileInput(
    inputId = "file",
    label = "Upload dataset",
    accept = c("csv")
  ),
  textInput(
    inputId = "name",
    label = "Dataset name"
  ),
  actionButton(
    inputId = "new",
    label = "Add to workspace",
    accept = c("csv")
  ),
  tags$hr(),
  tags$h4("Import/export workspace"),
  shiny::fileInput(
    inputId = "import",
    label = "Import workspace",
    accept = c("zip")
  ),
  downloadButton(
    outputId = "downloadData",
    label = "Export workspace"
  ),
  tags$hr(),
  tags$h4("Display workspace"),
  tableOutput("workspace"),
  selectInput(
    inputId = "selector",
    label = "Display table preview",
    choices = character()
  ),
  tableOutput("table")
)

server <- function(input, output) {
  rv <- reactiveValues(
    curr_dataset = NULL,
    ws = workspace::new_workspace(),
    summary = NULL
  )

  # Import csv file ----------
  observe({
    req(input$file)
    rv$curr_dataset <- shinyWidgets::execute_safely({
      readr::read_csv(input$file$datapath, show_col_types = FALSE)
    })
    updateTextInput(
      inputId = "name",
      value = tools::file_path_sans_ext(input$file$name)
    )
  })

  # Add new data set to workspace
  observeEvent(input$new, {
    dataset <- req(rv$curr_dataset)
    req(rv$ws)
    rv$ws <- workspace::store_dataset(x = rv$ws, dataset = dataset, name = trimws(input$name))
    rv$summary <- workspace::list_object_in_workspace(rv$ws)
  })

  # Display workspace metadata --------------
  output$workspace <- renderTable({
    req(rv$summary)
    rv$summary
  })

  observe({
    req(nrow(rv$summary) > 0)
    updateSelectInput(
      inputId = "selector",
      choices = rv$summary$name
    )
  })

  output$table <- renderTable({
    req(input$selector)
    workspace::read_dataset_in_workspace(x = rv$ws, name = input$selector) |>
      head()
  })

  # Export workspace ------------
  output$downloadData <- downloadHandler(
    filename = function() {
      paste("workspace-", Sys.Date(), ".zip", sep = "")
    },
    content = function(file) {
      workspace::pack_workspace(x = rv$ws, file = file)
    }
  )

  # Import workspace ----------
  observeEvent(input$import, {
    rv$ws <- workspace::unpack_workspace(file = input$import$datapath)
    rv$summary <- workspace::list_object_in_workspace(rv$ws)
  })
}

shinyApp(ui = ui, server = server)

```

