---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# workspace

`workspace` is an R package designed to simplify the creation and management of workspaces. Its main goal is to provide a standardized solution for organizing and sharing data in applications that require workspace management.

This package is aimed at developers and analysts seeking to standardize data management in multi-environment or cross-application contexts. With `workspace`, creating, manipulating, and sharing organized archives becomes straightforward.

## Key features

- Generation of structured archives:
  - Supports adding datasets in parquet format.
  - Enables storing JSON files for semi-structured data.
  - Handles R objects stored in RDS format.

- Enhanced interoperability:
  - Provides a well-defined and consistent structure for organizing data.
  - Facilitates data exchange and processing across different applications or environments.

- Offers tools to easily read and store data within workspaces.


## Installation

You can install the development version of workspace from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("ardata-fr/workspace")
```

## Example of workspace creation

This is a basic example which shows you how to create a workspace:

```{r}
library(workspace)
z <- new_workspace()

# store datasets
z <- store_dataset(x = z, dataset = iris, name = "iris_dataset")
z <- store_dataset(x = z, dataset = mtcars, name = "mtcars")

# store json
json_str <- paste0("{\"first_name\": \"John\",\"last_name\": \"Smith\",\"is_alive\": true,",
"\"age\": 27, \"address\": { \"street_address\": \"21 2nd Street\",",
"\"city\": \"New York\",\"state\": \"NY\",\"postal_code\": \"10021-3100\"",
"}}")
z <- store_json(
  x = z,
  name = "json-example",
  json_str = json_str,
  filename = "json-example.json",
  timestamp = "2023-11-12 11:37:41",
  subdir = "blah"
)

# pack workspace as a zip to share it
workspace_zip_file <- tempfile(fileext = ".zip")
pack_workspace(x = z, file = workspace_zip_file)
```

## Example of workspace data extraction

This is a basic example which shows you how to extract data from a workspace:

```{r}
z <- unpack_workspace(file = workspace_zip_file)
list_object_in_workspace(z)

dataset <- read_dataset_in_workspace(z, name = "mtcars")
print(head(dataset))

# store json
json_str <- read_json_str_in_workspace(z, name = "json-example", subdir = "blah")
substr(json_str, 1, 20) |> print()
```


